---
title: "HW 1"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---


Question 1
```{r}

su <- read.delim("C:/Users/vpool/Documents/Winter2024/CSC587DataMining/datasets/DATA/Su_raw_matrix.txt")
mean(su$Liver_2.CEL)
sd(su$Liver_2.cel)
colMeans(su)
colSums(su)
```

question 2.
```{r}
a <- rnorm(1000, mean = 0, sd = 0.2)
hist(a, main = "Histogram of a", xlim = c(-5,5)) 
b <-rnorm(1000, mean = 0, sd = 0.5)
hist(b, main = "Histogram of b", xlim = c(-5,5)) 
```
Histogram of a has have a narrower spread of data around the mean due to a smaller standard deviation 
vs b will have a wider spread of data due to a larger standard deviation meaning more varability of data 

Question 3.
```{r}
library (ggplot2)
dat <- data.frame(cond = factor(rep(c("A","B"), each = 200)), rating = c (rnorm(200, mean =.8)))
# OverLoad Histograms
ggplot(dat,aes(x=rating, fill = cond)) + geom_histogram(binwidth = .5, alpha = .5, position = "identity")
# Interleaved histogram
ggplot(dat, aes(x=rating, colour = cond)) + geom_histogram(binwidth = .5, position = "dodge")
#density plots 
ggplot(dat, aes(x=rating, colour = cond )) + geom_density()
# density plots with semitranspartent fill 
ggplot(dat, aes(x=rating, fill = cond)) + geom_density(alpha = .3)


diabetes <- read.csv("C:/Users/vpool/Documents/Winter2024/CSC587DataMining/datasets/DATA/diabetes_train.csv")
# OverLoad Histograms
ggplot(diabetes,aes(x=mass, fill = class)) + geom_histogram(binwidth = .5, alpha = .5, position = "identity")
# Interleaved histogram
ggplot(diabetes, aes(x=mass, colour = class)) + geom_histogram(binwidth = .5, position = "dodge")
#density plots 
ggplot(diabetes, aes(x=mass, colour = class )) + geom_density()
# density plots with semitranspartent fill 
ggplot(diabetes, aes(x=mass, fill = class)) + geom_density(alpha = .3)


```
questions 4 Read the titanic.csv file from DATA folder to a variable named passengers and perform the following steps and ex-
plain the operation very briefly. Please make sure you have tidyverse installed on your system and you may specifically
need to import the tidyr library. Otherwise, the chain of operations through ”piping” won’t work. (20 points):
(a) passengers %>% drop_na() %>% summary()
(b) passengers %>% filter(Sex == "male")

```{r}
library (tidyverse)
passengers <- read.csv("C:/Users/vpool/Documents/Winter2024/CSC587DataMining/datasets/DATA/titanic.csv")
passengers %>% drop_na() %>% summary()
passengers %>% filter(Sex == "male")
passengers %>% arrange(desc(Fare))
passengers %>% mutate (Famsize = Parch + SibSp)
passengers %>% group_by(Sex) %>% summarise(meanFare = mean(Fare), numSurv = sum(Survived)) 

```
Question 5 By using quantile(), calculate 10th, 30th, 50th, 60th percentiles of skin attribute of diabetes data. 
```{r}
quantile(diabetes$skin, probs = c(0.1, .3,.5,.6))


```